---
title: "Tillage meta-analysis"
author: "Dr. Chih-Yu Hung"
date: "2024-08-06"
output: html_document
---

## Objective

This document will analyze SOC changes due to tillage management in the USA. 

We have SOC changes (Mg C ha-1 y-1) in the US from publications. We need to prove that these empirical data matched the results from 


```{r preparation}
library(tidyverse)
library(ggplot2)
library(maps)
library(nlstools)

data.all <- read.csv("Input/SOC tillage_US.csv", header = T)
#45 studies
data.all.30 <- read.csv("Input/SOC tillage_US_30cm.csv",header =T )

data <- data.all %>%
  filter(Duration >= 3) %>%
  filter(Crop.type == "Annual") %>%
  filter(CT >0) %>%
  filter(SOC_Depth >=150)
#sort(unique(data$Article.title))   #40 studies, 244 obs

data.30 <- data.all.30 %>%
  filter(Duration >= 3) %>%
  filter(Crop.type == "Annual") %>%
  filter(CT >0)


#Convert soil texture to three texture system
data <- data %>%
  mutate(Soil.texture =
  case_when(tolower(str_trim(Texture)) %in% c("clay", "silty clay", "silty clay loam", "clay loam", "sandy clay") ~ "Fine",
            tolower(str_trim(Texture)) %in% c("sandy loam","sandy clay loam","loam soil", "loam", "silt loam", "silt") ~ "Medium",
            TRUE ~ "Coarse")
  )

data.30 <- data.30 %>%
  mutate(Soil.texture =
  case_when(tolower(str_trim(Texture)) %in% c("clay", "silty clay", "silty clay loam", "clay loam", "sandy clay") ~ "Fine",
            tolower(str_trim(Texture)) %in% c("sandy loam","sandy clay loam","loam soil", "loam", "silt loam", "silt") ~ "Medium",
            TRUE ~ "Coarse")
  )


#See the texture numbers
table(data$Soil.texture) #Fine 102, Medium 142 
table(data.30$Soil.texture) #Fine 57, Medium 128

# Filter data, separate to two groups North and South and group them to three duration, short, medium and long
data.N <- data %>%
  filter(Region == "North")%>%
  mutate(Period = case_when(
    Duration <= 10 ~ "Short",
    Duration >10 & Duration <=20 ~ "Mideum",
    Duration > 20 ~ "Long"))

#See the texture number in North states
table(data.N$Soil.texture) #Fine 79, Medium 100

data.S <- data %>%
  filter(Region == "South")%>%
  mutate(Period = case_when(
    Duration <= 10 ~ "Short",
    Duration >10 & Duration <=20 ~ "Mideum",
    Duration > 20 ~ "Long"))
#See the texture number
table(data.S$Soil.texture) #Fine 23, Medium 42


#Filter data to soil texture
data.Fine <- data %>%
  filter(Soil.texture == "Fine") %>%
  filter(Change_rate >= 0)

data.Medium <- data %>%
  filter(Soil.texture == "Medium") %>%
  filter(Change_rate >= 0)


```

#To see the clay content and the SOC changes for the coarse texture part. 



```{r US map to see data distribution}
# Assuming your dataframe is named "data" with a column "State"
# Count the number of occurrences per state
state_counts <- data %>%
  group_by(State) %>%
  summarize(Count = n())

# Get U.S. state map data
us_states <- map_data("state")

# Make sure the State names in your data match the map data format
state_counts$State <- tolower(state_counts$State)

# Calculate the centroids for each state to get long and lat
state_centroids <- us_states %>%
  group_by(region) %>%
  summarize(long = mean(range(long)), lat = mean(range(lat)))

# Merge the state centroids with the counts
state_counts <- state_counts %>%
  inner_join(state_centroids, by = c("State" = "region"))

# Plot the map
ggplot() +
  # Draw the U.S. state map
  geom_polygon(data = us_states, aes(x = long, y = lat, group = group),
               fill = "lightgray", color = "white") +
  # Add bubbles sized by the count
  geom_point(data = state_counts, aes(x = long, y = lat, size = Count),
             color = "blue", alpha = 0.6) +
  scale_size(range = c(1, 10), name = "Data Count") + # Adjust the range for bubble sizes
  labs(title = "Data Counts by State",
       x = "Longitude", y = "Latitude") +
  theme_minimal()

#To check data availablilty in the West
# data.west<- data %>%
#   filter(State %in% c("Montana", "Texas", "North Dakota")) %>%
#   filter(Soil.texture == "Medium")
```

find the statistic for production/area to see I cover 80% 

Repeated measurement and non-repeated measurement





## To check the average SOC change values in each durations

```{r group data}

average_north <- data.N %>%
  group_by(Period) %>%
  summarise(
    avg_Duration = mean(Duration),
    avg_Change_rate = mean(Change_rate)
  )
  

average_south <- data.S %>%
  group_by(Period) %>%
  summarise(
    avg_Duration = mean(Duration),
    avg_Change_rate = mean(Change_rate)
  )
  
```

```{r linear regression}

ggplot(data, aes(x = CT, y = NT)) +
  geom_point() +  # Plot the data points
  geom_smooth(method = "lm", formula = y ~ x + 0, color = "blue", se = TRUE) +  # Linear regression without intercept
  labs(
    title = "Linear Regression of NT vs CT (No Intercept)",
    x = "CT",
    y = "NT"
  ) +
  stat_poly_eq(
    formula = y ~ x + 0,
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    parse = TRUE,
    coef.digits = 3
  ) +
  theme_minimal()
```

```{r plot data}
ggplot(data.N, aes(x=Duration,y=Change_rate, color = Period)) +
  geom_point() +
  ggtitle("Northern US") +
  geom_point(data = average_north, aes(x = avg_Duration, y = avg_Change_rate), color = "black", size = 4, shape = 17) +  # Add average points
  xlab("Duration") +
  ylab("SOC change (kg C /ha /y)") +
  theme_classic()
  
  

ggplot(data.S, aes(x=Duration,y=Change_rate, color =Period)) +
  geom_point() +
  ggtitle("Southern US") +
  geom_point(data = average_south, aes(x = avg_Duration, y = avg_Change_rate), color = "black", size = 4, shape = 17) +  # Add average points
  xlab("Duration") +
  ylab("SOC change (kg C /ha /y)") +
  theme_classic()
  



```




```{r fit exponential decay models}

plot_exponential_decay <- function(data, max_C = 100, rate = 0.02) {
  library(ggplot2)
  library(nlstools)
  y_col <- "Change_rate"
  t_col <- "Duration"
  # Fit the model using nls with default start values if not provided
  fit <- nls(as.formula(paste(y_col, "~ a * exp(-k *", t_col, ")")),
             data = data, start = list(a = max_C, k = rate))
  
  # Extract the coefficients
  a <- coef(fit)["a"]
  k <- coef(fit)["k"]
  
  # Generate predictions and confidence intervals
  prediction_data <- data.frame(t_col = seq(min(data[[t_col]]), max(data[[t_col]]), length.out = 100))
  colnames(prediction_data)[1] <- t_col  # Set dynamic column name
  prediction_data$predicted <- predict(fit, newdata = prediction_data)
  
  # Get parameter confidence intervals
  conf <- confint2(fit, level = 0.95)
  a_upper <- conf["a", "97.5 %"]
  a_lower <- conf["a", "2.5 %"]
  k_upper <- conf["k", "97.5 %"]
  k_lower <- conf["k", "2.5 %"]
  
  # Calculate upper and lower bounds for confidence interval
  prediction_data$upper <- a_upper * exp(-k_lower * prediction_data[[t_col]])
  prediction_data$lower <- a_lower * exp(-k_upper * prediction_data[[t_col]])
  
  # Calculate RMSE and sMAPE
  data$predicted <- predict(fit)
  rmse <- sqrt(mean((data[[y_col]] - data$predicted)^2))
  sMAPE <- mean(2 * abs(data[[y_col]] - data$predicted) / (abs(data[[y_col]]) + abs(data$predicted))) * 100
  
  # Create labels for equation and metrics
  equation_label <- paste0("y = ", round(a, 3), " * exp(-", round(k, 3), " * t)")
  rmse_label <- paste0("RMSE = ", round(rmse, 3))
  sMAPE_label <- paste0("sMAPE = ", round(sMAPE, 2), "%")
  
# Plot with confidence interval and labels
  ggplot(data, aes_string(x = t_col, y = y_col)) +
    geom_point() +  # Plot data points
    geom_line(data = prediction_data, aes_string(x = t_col, y = "predicted"), color = "blue") +  # Fitted curve
   # geom_ribbon(data = prediction_data, aes_string(x = t_col, ymin = "lower", ymax = "upper"), alpha = 0.2) +  # Confidence interval
    labs(
      title = "Exponential Decay Fit with 95% Confidence Interval",
      x = t_col,
      y = y_col
    ) +
    annotate("text", x = Inf, y = Inf, label = equation_label, hjust = 1.1, vjust = 2, size = 4) +
    annotate("text", x = Inf, y = Inf, label = rmse_label, hjust = 1.1, vjust = 3.5, size = 4) +
    annotate("text", x = Inf, y = Inf, label = sMAPE_label, hjust = 1.1, vjust = 5, size = 4) +
    theme_minimal()
}

# Usage example:
# plot_exponential_decay(data, y_col = "Change_rate", t_col = "Duration")

plot

plot_exponential_decay(data.30)

```

