---
title: "SOC change due to tillage (CA NIR)"
author: "Dr. Chih-Yu Hung"
date: "2024-10-16"
output: html_document
---

## Purpose

We want to see the SOC changes in various Canada regions and soil textures. There are 9 regions and 3 soil textures. 

```{r setup}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(writexl)

CA_factor <- read.csv("Input/SOC tillage factor_CA.csv")
```

## Build a list for the total SOC from year 1 to year 100 (after practice change) in the 9 regions and 3 soil textures.

### Unit for deltaSOC, g C m-2, need to x 10 kg g-1 m-2 ha-1 to be kg C ha-1 
```{r calculate SOC change}

# Function to calculate SOC
calculate_SOC <- function(LUMCMAX, Kval, year) {
  SOC <- LUMCMAX * (1 - exp(-Kval * year))
  return(SOC)
}

# Create a list to store results
results <- list()

# Calculate SOC for each Zone, Texture, and year
for (i in 1:nrow(CA_factor)) {
  zone <- CA_factor$Zone[i]
  texture <- CA_factor$Texture[i]
  Kval <- CA_factor$Kval[i]
  LUMCMAX <- CA_factor$LUMCMAX[i]
  
  # Create a data frame for SOC over the years 1 to 100
  SOC_values <- data.frame(
    Year = 1:100,
    SOC = sapply(1:100, function(year) calculate_SOC(LUMCMAX, Kval, year))
  )
  
  # Store the SOC values in the list with Zone and Texture as keys
  results[[paste(zone, texture, sep = "_")]] <- SOC_values
}

# Display the results for inspection
results 

```



## Plot the SOC increase after turn to non-till in Canada 

```{r plot}
# Combine all the results into a single data frame for plotting
plot_data <- do.call(rbind, lapply(names(results), function(name) {
  df <- results[[name]]
  df$Combination <- name
  return(df)
}))

# Create the plot
ggplot(plot_data, aes(x = Year, y = SOC, color = Combination, linetype = Combination)) +
  geom_line(size = 1) +
  labs(
    title = "SOC over Time for Di
    fferent Zone and Texture Combinations",
    x = "Year",
    y = "SOC",
    color = "Zone_Texture Combination",
    linetype = "Zone_Texture Combination"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  ) +
  scale_color_viridis_d() +
  scale_linetype_manual(values = rep(1:6, length.out = 33))



#Seperated by soil texture
# Combine all the results into a single data frame for plotting, including Texture information
plot_data_texture <- do.call(rbind, lapply(1:nrow(CA_factor), function(i) {
  df <- results[[paste(CA_factor$Zone[i], CA_factor$Texture[i], sep = "_")]]
  df$Zone <- CA_factor$Zone[i]
  df$Texture <- CA_factor$Texture[i]
  return(df)
}))

#Plot by texture
plot_texture <- ggplot(plot_data_texture, aes(x = Year, y = SOC, color = Zone, linetype = Zone)) +
  geom_line(size = 1.2) +  # Increase the line size for better visibility
  labs(
    title = "SOC over Time for Different Zone and Texture Combinations",
    x = "Year",
    y = "SOC",
    color = "Zone",
    linetype = "Zone"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  ) +
  scale_color_manual(values = RColorBrewer::brewer.pal(11, "Set3")) +  # Use a palette with 11 distinct colors
  facet_wrap(~Texture, ncol = 1, scales = "free_y") +
  scale_linetype_manual(values = c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash", 
                                   "solid", "dashed", "dotted", "dotdash", "longdash"))  # 11 line types

ggsave("Output/SOC_over_Time_by_Texture.png", plot = plot_texture, width = 10, height = 12, dpi = 300)
```



## Calculate Avg. delta SOC to compare my results from US data

```{r annual SOC change}
#This results should be comparable to my meta analysis in US

deltaSOC <- lapply(results, function(df) {
 # This is the average deltaSOC for the given years
  df$deltaSOC_Avg <- c(df$SOC*10/df$Year) #convert to kg C ha-1
  return(df)
})

save(deltaSOC, file="Output/deltaSOC_CA.RData")

#write_xlsx(deltaSOC, path = "Output/DeltaSOC_CA.xlsx")


#To see the avg for each period
calculate_period_avgs <- function(df) {
  df <- df %>%
    filter(Year >= 3 & Year <= 50) %>%  # Keep only rows within the range
    mutate(
      Period = case_when(
        Year >= 3 & Year <= 10 ~ "short",
        Year >= 11 & Year <= 20 ~ "medium",
        Year >= 21 & Year <= 50 ~ "long",
      )
    ) %>%
    group_by(Period) %>%
    summarise(avg_deltaSOC_Avg = round(mean(deltaSOC_Avg, na.rm = TRUE),1), .groups = "drop")
  
  return(df)
}

# Apply the function to each data frame in the list
results_list <- lapply(deltaSOC, calculate_period_avgs)

# Combine all results into a single summary table
final_results <- bind_rows(results_list, .id = "DataFrame")


# Pivot the table to have Period as columns
pivoted_results <- final_results %>%
  pivot_wider(
    names_from = Period,
    values_from = avg_deltaSOC_Avg,
    values_fill = NA  # Fill missing combinations with NA
  )

# Inspect the final results
print(final_results)





```


```{r relationship between coarse and fine soils}
#To see if I can find a adjustment factor for Coarse texture soils
# Function to add the factor columns and return a new list with only "_Coarse" data frames
add_factors_and_filter <- function(data_list) {
  # Create an empty list to store modified data frames with "_Coarse"
  modified_data_list <- list()
  
  # Iterate over each dataframe in the list
  for (df_name in names(data_list)) {
    # Check if the data frame name ends with "_Coarse"
    if (grepl("_Coarse$", df_name)) {
      # Extract the base name (e.g., "Boreal Shield West")
      base_name <- sub("_Coarse$", "", df_name)
      
      # Define the related data frames for Medium and Fine textures
      medium_df_name <- paste0(base_name, "_Medium")
      fine_df_name <- paste0(base_name, "_Fine")
      
      # Check if both the Medium and Fine data frames exist in the list
      if (medium_df_name %in% names(data_list) && fine_df_name %in% names(data_list)) {
        # Add factor_Medium and factor_Fine columns to the "_Coarse" dataframe
        data_list[[df_name]]$factor_Medium <- data_list[[df_name]]$deltaSOC / data_list[[medium_df_name]]$deltaSOC
        data_list[[df_name]]$factor_Fine <- data_list[[df_name]]$deltaSOC / data_list[[fine_df_name]]$deltaSOC
        
        # Store the modified "_Coarse" dataframe in the new list
        modified_data_list[[df_name]] <- data_list[[df_name]]
      }
    }
  }
  
  # Return the new list containing only "_Coarse" data frames with factors
  return(modified_data_list)
}

# Apply the function to your list
filtered_data_list <- add_factors_and_filter(deltaSOC)

# Apply the function to your list and save it back to data_list
data_list <- add_factors(deltaSOC)



```